### 2017-08-16 Tex_Dissertation/MSc_dissertation.tex
alm

@@

The point that particles collide in a collider experiment is called primary vertex or interaction 
point (IP). After that 
collision, it is possible that some of the particles will decay further away from the IP. The point 
that this decay takes place is called secondary vertex. Every analysis that involves jet clustering
procedure (and not only) has as starting point the vertex finding. This involves an algorithm

For the needs of this projecct I used the package of LCFIplus which is an improved software of the already 
existing one LCFIVertex. The important difference between the  two is that LCFIplus employs the vertex finding 
before the jet clustering (in comparison to LCFIVertex which follows the exact opposite procedure) with the 
goal to protect the secondadry vertices from breaking up and grouped into different jets~\cite{suehara2016lcfiplus}.
This is an important step towards better performance since the Bottom jets can be identified from their 
sufficiently large distance between the primary vertex and their decay vertex. Additionally, the importance
of this step to be performed before jet clustering, is that hadrons that contain bottom or charm quarks 
have sizeable lifetimes and they can be measured inside the detector~\cite{suehara2016lcfiplus}.

@@

Jets are an indespensible part of high energy physics. They come from the manifestation of strong nuclear
as free quarks or gluons hadronize. They are visual structures that are measured in the detectors and they
serve the need of approaching the initial parton that they originated. Throught the years various algorithms 
have been developed for the clustering of such objects differing in the way that they approach such procedure.
Two of the most famous are the sequential recombination algorithms and cone algorithms. In this projecct I use
the a sequential recombination algorithm, the $k_{t}$ algorithm which is a longitudinaly invariant algorithm
firstly designed for hadron colliders.

@@

where $p_{Ti}$ is the transverse momentum of $i^{th}$ particle with respect to beam direction and 
$\Delta R^{2}_{ij}=(y_{i}-y_{j})^{2}+(\phi_{i}-\phi_{j})^{2}$ with $y$ being the rapitdity, $\phi$ being
the azimuth angle and R the jet radius fixed in this project as $R=0.7$.

The basic idea behind this algorithm is that it calculates the $d_{ij}$ with the smallest value and then it 
"merges" the $i,j$ into a single object with momentum $p_{i}+p_{j}$. This procedure is repeated until the quantity
$d_{ij}$ is above some threshold and all the particles that are left are the event's jet
~\cite{cacciari2012fastjet}.

\subsection{Flavour Tagging}

Flavour Tagging is the procedure in which a pseudoprobability is assigned to each jet regarding the parton
that the jet originates from. In this project, there are six jets in the final products four of them 
originating from b quarks.

The flavour tagging algorithm that was used belonged in the LCFIplus package~\cite{suehara2016lcfiplus}.
It  makes use of the TMVA package which in turn uses multivariate classifiers. A set of input variables are 
defined for each jet and they are passed to the Gradien Boosted Descision Trees (a certain type of classifiers)
which they are trained and then applied to the jets of the physical process. Those input variables are
constituents of the jets such as tracks and secondary vertices.

>>>>

 In this project I investigate the first channel with the W boson decaying into hadrons. This results
 into having 6 jets in the final products. The channel explicitly is:

 $e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t \bar{t}\tilde{\chi}_{1}^{0}
 \tilde{\chi}^{0}_{1} \rightarrow
 W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0} \rightarrow 
 q \bar{q} q \bar{q} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}$

@@

The point that particles collide in a collider experiment is called primary vertex or interaction 
point (IP). After the 
collision, it is possible that some of the particles will decay further away from the IP. The point 
that this decay takes place is called secondary vertex. Every analysis that involves jet clustering
procedure (and not only) has as starting point the vertex finding. This involves an algorithm that 
classifies the tracks of the particles into vertex candidates (vertex finding) and "passes it on"
to the vertex fitting  algorithm. In turn, the latter, takes the vertex candidates and outputs a
list of vertices with their position using the $\chi^{2}$ or a related statistical method. 
Additionally, it yields a set of parameters that are related to each particular production point.

The reasons of why an accurate vertex finding is important are:

\begin{itemize}
 \item Information can be exctracted about short lived particles such as their decay length.
 \item There can be improvement with respect to momentum estimate of the particles involved in each 
 vertex.
 \item Reconstruction can be achieved for short lived particles by its decay products by fitting them
 into one particular vertex.
\end{itemize}

For the needs of this projecct I used the package of LCFIplus which is an improved software of the already 
existing one LCFIVertex. The important difference between the  two is that LCFIplus employs the vertex finding 
before the jet clustering (in comparison to LCFIVertex which follows the exact opposite procedure) with the 
goal to protect the secondary vertices from breaking up and grouped into different jets~\cite{suehara2016lcfiplus}.
This is an important step towards better performance since the bottom jets can be identified from their 
sufficiently large distance between the primary vertex and their decay vertex. Additionally, the importance
of this step to be performed before jet clustering, is that hadrons that contain bottom or charm quarks 
have sizeable lifetimes and they can be measured inside the detector~\cite{suehara2016lcfiplus}.

@@

Jets are an indespensible part of high energy physics. They come from the manifestation of strong nuclear
as free quarks or gluons hadronize. They are visual structures that are measured in the detectors and they
serve the need of approaching the initial parton that they originated. Throught the years various algorithms 
have been developed for the clustering of such objects differing in the way that they approach such procedure.
Two of the most famous are the sequential recombination algorithms and cone algorithms. In this projecct I use
the a sequential recombination algorithm, the $k_{t}$ algorithm which is a longitudinaly invariant algorithm
firstly designed for hadron colliders~\cite{catani1993longitudinally}~\cite{ellis1993successive}.

@@

where $p_{Ti}$ is the transverse momentum of $i^{th}$ particle with respect to beam direction and 
$\Delta R^{2}_{ij}=(y_{i}-y_{j})^{2}+(\phi_{i}-\phi_{j})^{2}$ with $y$ being the rapitdity, $\phi$ being
the azimuth angle and $R$ the jet radius fixed in this project as $R=0.7$.

The basic idea behind this algorithm is that it calculates the $d_{ij}$ with the smallest value and then it 
"merges" the $i,j$ into a single object with momentum $p_{i}+p_{j}$. This procedure is repeated until the quantity
$d_{ij}$ is above some threshold and all the particles that are left are the event's jet~\cite{cacciari2012fastjet}.

For this project the jet clustering procedure is stoped when the number of jets reaches the number 
6, since in the final products there are 6 jets (see section 4.1).
\subsection{Flavour Tagging}

Flavour Tagging is the procedure in which a pseudoprobability is assigned to each jet regarding the
flavour of the parton
that the jet originates from. In this project, there are six jets in the final products two of them 
originating from b quarks.

The flavour tagging algorithm belongs in the LCFIplus package~\cite{suehara2016lcfiplus}.
It  makes use of the TMVA package which in turn uses multivariate classifiers. A set of input variables are 
defined for each jet and they are passed to the Gradient Boosted Descision Trees (a certain type of classifiers)
which they are trained and then applied to the jets of the physical process. Those input variables are
constituents of the jets such as tracks and secondary vertices.
